---
title: "COVID-19 Analyses"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)

library(caret)
library(pROC)
```

##  Oversampling and undersampling exploration for MixModel{.tabset}

Independet variables:

* `PT`
* `Neutrophilspercent`
* `lymphocytespercent`
* `CatHaematocrit`
* `Age`
* `CRP`
* `Respiratorydistress1`
* `ASTALTRatio`
* `APTT`
* `BloodUreaNitrogen`

```{r full data loading, echo=FALSE, cache=TRUE}
Farvadin_2 <- read_dta("fwdcovid19/15 Farvardin-2 dleted missing discharge date.dta", encoding = "windows-1256") %>% 
  mutate (Outcome = as.factor(Outcome)) %>%
  select(Outcome,
         PT,
         Neutrophilspercent,
         lymphocytespercent,
         CatHaematocrit,
         Age,
         CRP,
         Respiratorydistress1,
         ASTALTRatio,
         APTT,
         BloodUreaNitrogen) %>% 
  # filter(complete.cases(.)) %>% 
  mutate(Outcome = recode(Outcome, "0" = "ALIVE", "1" = "DEATH")) %>% 
  fastDummies::dummy_columns(select_columns = c("CatHaematocrit", "CatHB"), 
                             ignore_na = T,
                             remove_selected_columns = T) -> Farvadin_2_full_data
```


### Reference model

Train and test datasets created previoysly using `createDataPartition(as.factor(Farvadin_2$Outcome), p=0.80, list=FALSE)` 

Train data (80%)

```{r train data loading, echo=FALSE, cache = TRUE}
dtrain_80 <- readRDS("Farvadin_2_train_80.Rda")  %>% 
  mutate(Outcome = recode(Outcome, "0" = "ALIVE", "1" = "DEATH"))
summary(dtrain_80$Outcome)
```

Test data (20%)

```{r test data loading, echo=FALSE, warning = F, cache = TRUE}
dtest_20 <- readRDS("Farvadin_2_test_20.Rda") %>% 
  mutate(Outcome = recode(Outcome, "0" = "ALIVE", "1" = "DEATH"))
summary(dtest_20$Outcome)
```

#### Loading data

```{r mixmodel load  data, echo=FALSE, cache = TRUE}
dtrain_80 %>%
  select(Outcome,
         PT,
         Neutrophilspercent,
         lymphocytespercent,
         CatHaematocrit,
         Age,
         CRP,
         Respiratorydistress1,
         ASTALTRatio,
         APTT,
         BloodUreaNitrogen) %>% 
  # filter(complete.cases(.)) %>% 
  fastDummies::dummy_columns(select_columns = c("CatHaematocrit", "CatHB"), 
                             ignore_na = T,
                             remove_selected_columns = T) -> mixmodel_dtrain_80

dtest_20 %>%
  select(Outcome,
         PT,
         Neutrophilspercent,
         lymphocytespercent,
         CatHaematocrit,
         Age,
         CRP,
         Respiratorydistress1,
         ASTALTRatio,
         APTT,
         BloodUreaNitrogen) %>% 
  # filter(complete.cases(.)) %>% 
  fastDummies::dummy_columns(select_columns = c("CatHaematocrit", "CatHB"), 
                             ignore_na = T,
                             remove_selected_columns = T) -> mixmodel_dtest_20

```

#### Model trains

```{r mixmodel model train, echo=FALSE, cache = TRUE}
library(doParallel)
cl <- makePSOCKcluster(detectCores())
registerDoParallel(cl)
mixmodel_train_control = trainControl(
  method = "cv",
  number = 5,
  verboseIter = FALSE,
  returnData = FALSE,
  returnResamp = "all", # save losses across all models
  classProbs = TRUE, # set to TRUE for AUC to be computed
  summaryFunction = twoClassSummary,
  allowParallel = TRUE
)

mixmodel_tune_grid = expand.grid(
  nrounds = 5000,
  max_depth = c(2, 4, 6, 8, 10),
  eta=c(0.5, 0.3, 0.1, 0.07),
  gamma = c(0, 0.01),
  colsample_bytree=c(1, 0.5),
  min_child_weight=1,
  subsample=c(1, 0.5)
)

# Recode https://stackoverflow.com/a/44131491/9664743
set.seed(0)
mixmodel_caret_model = train(
  x = as.matrix(mixmodel_dtrain_80 %>% select(-Outcome)),
  y = mixmodel_dtrain_80$Outcome,
  trControl = mixmodel_train_control,
  tuneGrid = mixmodel_tune_grid,
  metric= "ROC",
  method = "xgbTree"
)
registerDoSEQ()
mixmodel_caret_model$bestTune
```

```{r mixmodel care model save, echo=FALSE}
saveRDS(mixmodel_caret_model, file="mixmodel_caret_train_20121204.model")
```

##### Variables importance

```{r mixmodel variable importance, echo=FALSE, cache=T}
mixmodel_var_importance <- varImp(mixmodel_caret_model, scale = FALSE, dig=3)
mixmodel_var_importance
```


```{r mixmodel variable importance plot, echo=FALSE, cache=T}

## BloodUreaNitrogen    0.19269
## ASTALTRatio          0.16302
## Neutrophilspercent   0.13521
## APTT                 0.13216
## CRP                  0.11195
## Age                  0.09208
## lymphocytespercent   0.07778
## PT                   0.07214
## Respiratorydistress1 0.00934
## CatHaematocrit_2     0.00747
## CatHaematocrit_1     0.00340
## CatHaematocrit_0     0.00275

rownames(mixmodel_var_importance$importance) <- 
  c("Blood Urea Nitrogen",
    "AST/ALT Ratio",
    "Neutrophils % ",
    "APTT",
    "CRP",
    "Age",
    "Lymphocytes %",
    "PT",
    "Respiratory distress",
    "Haematocrit - More",
    "Haematocrit - Normal",
    "Haematocrit - Less")
# plot(mixmodel_var_importance)
ggplot(mixmodel_var_importance)
```

#### Model results

##### AUC

```{r mixmodel auc, warning=T, echo=FALSE, cache = T}
mixmodel_pred_prob <- predict(mixmodel_caret_model,
                                as.matrix(mixmodel_dtest_20 %>% select(-Outcome)), type="prob")
#plot((roc(response=mixmodel_dtest_20$Outcome, predictor=mixmodel_pred$DEATH)),print.thres="best", print.thres.best.method="closest.topleft")
roc(response=mixmodel_dtest_20$Outcome, predictor=mixmodel_pred_prob$DEATH, levels = c("ALIVE", "DEATH"), direction = c("<"))
```

##### Predictivy values
```{r mixmodel predictive values, echo=FALSE, cache = T}
mixmodel_pred <- predict(mixmodel_caret_model,
                           as.matrix(mixmodel_dtest_20 %>% select(-Outcome)))
confusionMatrix(data = mixmodel_pred,
                reference = mixmodel_dtest_20$Outcome,
                positive = "DEATH")
```

Confidence intervals of Sensitivity and Specificity (Note that Outcome + means 'DEATH')

```{r mixmodel predictive values w CIs, echo=FALSE, cache = T}
library(epiR)
options(digits = 3)
confusionMatrix(data = factor(mixmodel_pred,levels=c("DEATH", "ALIVE")) , reference = factor(mixmodel_dtest_20$Outcome, levels=c("DEATH", "ALIVE")), positive = "DEATH", mode=c("sens_spec"))$table -> mixmodel_matrix_table
epiR::epi.tests(mixmodel_matrix_table) -> mixmodel_epitests

Sensitivity <- c(mixmodel_epitests$elements$se, mixmodel_epitests$elements$se.low, mixmodel_epitests$elements$se.up)
Specificity <- c(mixmodel_epitests$elements$sp, mixmodel_epitests$elements$sp.low, mixmodel_epitests$elements$sp.up)
`Positive Predictive Value` <- c(mixmodel_epitests$elements$ppv, mixmodel_epitests$elements$ppv.low, mixmodel_epitests$elements$ppv.up)
`Negative Predictive Value` <- c(mixmodel_epitests$elements$npv, mixmodel_epitests$elements$npv.low, mixmodel_epitests$elements$npv.up)
mixmodel_epitests_df <- as.data.frame(rbind(Sensitivity, Specificity, `Positive Predictive Value`, `Negative Predictive Value`))
colnames(mixmodel_epitests_df) <- c("Estimate", "Lower Interval", "Upper Interval")
knitr::kable(mixmodel_epitests_df)
```

### Oversampled model

Train and test datasets created previoysly using `upSample` from `caret` package and the `createDataPartition(as.factor(Farvadin_2$Outcome), p=0.80, list=FALSE)` also from the `caret` package.

Train data (80%)

```{r oversampled train data loading, cache = TRUE}
dtrain_over_80 <- readRDS("Farvadin_2_balanced_over_train_80.Rda")  %>% 
  mutate(Outcome = recode(Outcome, "0" = "ALIVE", "1" = "DEATH"))
summary(dtrain_over_80$Outcome)
```

Test data (20%)

```{r oversampled test data loading, warning = F, cache = TRUE}
dtest_over_20 <- readRDS("Farvadin_2_balanced_over_test_20.Rda") %>% 
  mutate(Outcome = recode(Outcome, "0" = "ALIVE", "1" = "DEATH"))
summary(dtest_over_20$Outcome)
```

#### Loading data

```{r oversampled load  data, cache = TRUE}
dtrain_over_80 %>%
  select(Outcome,
         PT,
         Neutrophilspercent,
         lymphocytespercent,
         CatHaematocrit,
         Age,
         CRP,
         Respiratorydistress1,
         ASTALTRatio,
         APTT,
         BloodUreaNitrogen) %>% 
  # filter(complete.cases(.)) %>% 
  fastDummies::dummy_columns(select_columns = c("CatHaematocrit", "CatHB"), 
                             ignore_na = T,
                             remove_selected_columns = T) -> mixmodel_dtrain_over_80

dtest_over_20 %>%
  select(Outcome,
         PT,
         Neutrophilspercent,
         lymphocytespercent,
         CatHaematocrit,
         Age,
         CRP,
         Respiratorydistress1,
         ASTALTRatio,
         APTT,
         BloodUreaNitrogen) %>% 
  # filter(complete.cases(.)) %>% 
  fastDummies::dummy_columns(select_columns = c("CatHaematocrit", "CatHB"), 
                             ignore_na = T,
                             remove_selected_columns = T) -> mixmodel_dtest_over_20

```

#### Model trains

```{r oversampled model train, cache = TRUE}
library(doParallel)
cl <- makePSOCKcluster(detectCores())
registerDoParallel(cl)
mixmodel_over_train_control = trainControl(
  method = "cv",
  number = 5,
  verboseIter = FALSE,
  returnData = FALSE,
  returnResamp = "all", # save losses across all models
  classProbs = TRUE, # set to TRUE for AUC to be computed
  summaryFunction = twoClassSummary,
  allowParallel = TRUE
)

mixmodel_over_tune_grid = expand.grid(
  nrounds = 5000,
  max_depth = c(2, 4, 6, 8, 10),
  eta=c(0.5, 0.3, 0.1, 0.07),
  gamma = c(0, 0.01),
  colsample_bytree=c(1, 0.5),
  min_child_weight=1,
  subsample=c(1, 0.5)
)

# Recode https://stackoverflow.com/a/44131491/9664743
set.seed(0)
mixmodel_over_caret_model = train(
  x = as.matrix(mixmodel_dtrain_over_80 %>% select(-Outcome)),
  y = mixmodel_dtrain_over_80$Outcome,
  trControl = mixmodel_over_train_control,
  tuneGrid = mixmodel_over_tune_grid,
  metric= "ROC",
  method = "xgbTree"
)
registerDoSEQ()
mixmodel_over_caret_model$bestTune
```

```{r oversampled care model save}
saveRDS(mixmodel_over_caret_model, file="mixmodel_over_caret_train_20121204.model")
```

#### Variables importance

```{r oversampled variable importance full data, echo=FALSE, cache=T}
mixmodel_over_var_importance <- varImp(mixmodel_over_caret_model, scale = FALSE, dig=3)
mixmodel_over_var_importance
```


```{r oversampled variable importance plot f, echo=FALSE, cache=T}
## lymphocytespercent   0.155644
## BloodUreaNitrogen    0.144569
## APTT                 0.134912
## Neutrophilspercent   0.127320
## Age                  0.114644
## ASTALTRatio          0.113001
## CRP                  0.093424
## PT                   0.073857
## CatHaematocrit_0     0.016608
## CatHaematocrit_2     0.013890
## Respiratorydistress1 0.009435
## CatHaematocrit_1     0.002696
rownames(mixmodel_over_var_importance$importance) <- 
  c("Lymphocytes %",
    "Blood Urea Nitrogen",
    "APTT",
    "Neutrophils % ",
    "Age",
    "AST/ALT Ratio",
    "CRP",
    "PT",
    "Haematocrit - Less",
    "Haematocrit - More",
    "Respiratory distress",
    "Haematocrit - Normal")
ggplot(mixmodel_over_var_importance)
```

#### Model results using oversample TEST set 

##### AUC

```{r oversampled auc, warning=T, echo=FALSE, cache = T}
mixmodel_over_pred_prob <- predict(mixmodel_over_caret_model,
                                as.matrix(mixmodel_dtest_over_20 %>% select(-Outcome)), type="prob")
#plot((roc(response=mixmodel_dtest_20$Outcome, predictor=mixmodel_pred$DEATH)),print.thres="best", print.thres.best.method="closest.topleft")
roc(response=mixmodel_dtest_over_20$Outcome, predictor=mixmodel_over_pred_prob$DEATH, levels = c("ALIVE", "DEATH"), direction = c("<"))
```

##### Predictivy values
```{r oversampled predictive values, echo=FALSE, cache = T}
mixmodel_over_pred <- predict(mixmodel_over_caret_model,
                           as.matrix(mixmodel_dtest_over_20 %>% select(-Outcome)))
confusionMatrix(data = mixmodel_over_pred,
                reference = mixmodel_dtest_over_20$Outcome,
                positive = "DEATH")
```

Confidence intervals of Sensitivity and Specificity (Note that Outcome + means 'DEATH')

```{r oversampled predictive values w CIs, echo=FALSE, cache = T}
library(epiR)
options(digits = 3)
confusionMatrix(data = factor(mixmodel_over_pred,levels=c("DEATH", "ALIVE")) , reference = factor(mixmodel_dtest_over_20$Outcome, levels=c("DEATH", "ALIVE")), positive = "DEATH", mode=c("sens_spec"))$table -> mixmodel_over_matrix_table
epiR::epi.tests(mixmodel_over_matrix_table) -> mixmodel_over_epitests

Sensitivity <- c(mixmodel_over_epitests$elements$se, mixmodel_over_epitests$elements$se.low, mixmodel_over_epitests$elements$se.up)
Specificity <- c(mixmodel_over_epitests$elements$sp, mixmodel_over_epitests$elements$sp.low, mixmodel_over_epitests$elements$sp.up)
`Positive Predictive Value` <- c(mixmodel_over_epitests$elements$ppv, mixmodel_over_epitests$elements$ppv.low, mixmodel_over_epitests$elements$ppv.up)
`Negative Predictive Value` <- c(mixmodel_over_epitests$elements$npv, mixmodel_over_epitests$elements$npv.low, mixmodel_over_epitests$elements$npv.up)
mixmodel_epitests_df <- as.data.frame(rbind(Sensitivity, Specificity, `Positive Predictive Value`, `Negative Predictive Value`))
colnames(mixmodel_epitests_df) <- c("Estimate", "Lower Interval", "Upper Interval")
knitr::kable(mixmodel_epitests_df)
```

#### Model results using all data 

##### AUC

```{r oversampled auc full data, warning=T, echo=FALSE, cache = T}
mixmodel_over_full_data_pred_prob <- predict(mixmodel_over_caret_model,
                                             as.matrix(Farvadin_2_full_data %>% select(-Outcome)), type="prob")
#plot((roc(response=mixmodel_dtest_20$Outcome, predictor=mixmodel_pred$DEATH)),print.thres="best", print.thres.best.method="closest.topleft")
roc(response=Farvadin_2_full_data$Outcome, predictor=mixmodel_over_full_data_pred_prob$DEATH, levels = c("ALIVE", "DEATH"), direction = c("<"))
```

##### Predictivy values
```{r oversampled predictive values full data, echo=FALSE, cache = T}
mixmodel_over_full_data_pred <- predict(mixmodel_over_caret_model,
                                        as.matrix(Farvadin_2_full_data %>% select(-Outcome)))
confusionMatrix(data = mixmodel_over_full_data_pred,
                reference = Farvadin_2_full_data$Outcome,
                positive = "DEATH")
```

Confidence intervals of Sensitivity and Specificity (Note that Outcome + means 'DEATH')

```{r oversampled predictive values w CIs full data, echo=FALSE, cache = T}
library(epiR)
options(digits = 3)
confusionMatrix(data = factor(mixmodel_over_full_data_pred,levels=c("DEATH", "ALIVE")) ,
                reference = factor(Farvadin_2_full_data$Outcome, levels=c("DEATH", "ALIVE")),
                positive = "DEATH", mode=c("sens_spec"))$table -> mixmodel_over_full_data_matrix_table
epiR::epi.tests(mixmodel_over_full_data_matrix_table) -> mixmodel_over_full_data_epitests

Sensitivity <- c(mixmodel_over_full_data_epitests$elements$se, mixmodel_over_full_data_epitests$elements$se.low, mixmodel_over_full_data_epitests$elements$se.up)

Specificity <- c(mixmodel_over_full_data_epitests$elements$sp, mixmodel_over_full_data_epitests$elements$sp.low, mixmodel_over_full_data_epitests$elements$sp.up)

`Positive Predictive Value` <- c(mixmodel_over_full_data_epitests$elements$ppv, mixmodel_over_full_data_epitests$elements$ppv.low, mixmodel_over_full_data_epitests$elements$ppv.up)

`Negative Predictive Value` <- c(mixmodel_over_full_data_epitests$elements$npv, mixmodel_over_full_data_epitests$elements$npv.low, mixmodel_over_full_data_epitests$elements$npv.up)

mixmodel_epitests_full_data_df <- as.data.frame(rbind(Sensitivity, Specificity, `Positive Predictive Value`, `Negative Predictive Value`))
colnames(mixmodel_epitests_full_data_df) <- c("Estimate", "Lower Interval", "Upper Interval")

knitr::kable(mixmodel_epitests_full_data_df)
```

### Undersample model

Train and test datasets created previoysly using `downSample` from `caret` package and the `createDataPartition(as.factor(Farvadin_2$Outcome), p=0.80, list=FALSE)` also from the `caret` package.

Train data (80%)

```{r undersampled train data loading, echo=FALSE, cache = TRUE}
dtrain_under_80 <- readRDS("Farvadin_2_balanced_under_train_80.Rda")  %>% 
  mutate(Outcome = recode(Outcome, "0" = "ALIVE", "1" = "DEATH"))
summary(dtrain_under_80$Outcome)
```

Test data (20%)

```{r undersampled test data loading, echo=FALSE, warning = F, cache = TRUE}
dtest_under_20 <- readRDS("Farvadin_2_balanced_under_test_20.Rda") %>% 
  mutate(Outcome = recode(Outcome, "0" = "ALIVE", "1" = "DEATH"))
summary(dtest_under_20$Outcome)
```

#### Loading data

```{r undersampled load  data, echo=FALSE, cache = TRUE}
dtrain_under_80 %>%
  select(Outcome,
         PT,
         Neutrophilspercent,
         lymphocytespercent,
         CatHaematocrit,
         Age,
         CRP,
         Respiratorydistress1,
         ASTALTRatio,
         APTT,
         BloodUreaNitrogen) %>% 
  # filter(complete.cases(.)) %>% 
  fastDummies::dummy_columns(select_columns = c("CatHaematocrit", "CatHB"), 
                             ignore_na = T,
                             remove_selected_columns = T) -> mixmodel_dtrain_under_80

dtest_under_20 %>%
  select(Outcome,
         PT,
         Neutrophilspercent,
         lymphocytespercent,
         CatHaematocrit,
         Age,
         CRP,
         Respiratorydistress1,
         ASTALTRatio,
         APTT,
         BloodUreaNitrogen) %>% 
  # filter(complete.cases(.)) %>% 
  fastDummies::dummy_columns(select_columns = c("CatHaematocrit", "CatHB"), 
                             ignore_na = T,
                             remove_selected_columns = T) -> mixmodel_dtest_under_20

```

#### Model trains

```{r undersampled model train, echo=FALSE, cache = TRUE}
library(doParallel)
cl <- makePSOCKcluster(detectCores())
registerDoParallel(cl)
mixmodel_under_train_control = trainControl(
  method = "cv",
  number = 5,
  verboseIter = FALSE,
  returnData = FALSE,
  returnResamp = "all", # save losses across all models
  classProbs = TRUE, # set to TRUE for AUC to be computed
  summaryFunction = twoClassSummary,
  allowParallel = TRUE
)

mixmodel_under_tune_grid = expand.grid(
  nrounds = 5000,
  max_depth = c(2, 4, 6, 8, 10),
  eta=c(0.5, 0.3, 0.1, 0.07),
  gamma = c(0, 0.01),
  colsample_bytree=c(1, 0.5),
  min_child_weight=1,
  subsample=c(1, 0.5)
)

# Recode https://stackoverflow.com/a/44131491/9664743
set.seed(0)
mixmodel_under_caret_model = train(
  x = as.matrix(mixmodel_dtrain_under_80 %>% select(-Outcome)),
  y = mixmodel_dtrain_under_80$Outcome,
  trControl = mixmodel_under_train_control,
  tuneGrid = mixmodel_under_tune_grid,
  metric= "ROC",
  method = "xgbTree"
)
registerDoSEQ()
mixmodel_under_caret_model$bestTune
```

```{r undersampled care model save}
saveRDS(mixmodel_under_caret_model, file="mixmodel_under_caret_train_20121204.model")
```
##### Variables importance

```{r undersampled variable importance, echo=FALSE, cache=T}
mixmodel_under_var_importance <- varImp(mixmodel_under_caret_model, scale = FALSE, dig=3)
mixmodel_under_var_importance
```


```{r undersampled variable importance plot, echo=FALSE, cache=T}

## Neutrophilspercent   0.35931
## Age                  0.21193
## BloodUreaNitrogen    0.15784
## ASTALTRatio          0.09336
## APTT                 0.06541
## lymphocytespercent   0.04389
## CRP                  0.03440
## PT                   0.01741
## CatHaematocrit_0     0.01351
## Respiratorydistress1 0.00293
## CatHaematocrit_2     0.00000
## CatHaematocrit_1     0.00000

rownames(mixmodel_under_var_importance$importance) <- 
  c("Neutrophils % ",
    "Age",
    "Blood Urea Nitrogen",
    "AST/ALT Ratio",
    "APTT",
    "Lymphocytes %",
    "CRP",
    "PT",
    "Haematocrit - Less",
    "Respiratory distress",
    "Haematocrit - More",
    "Haematocrit - Normal")
# plot(mixmodel_var_importance)
ggplot(mixmodel_under_var_importance)
```

#### Model results using oversample TEST set

##### AUC

```{r undersampled auc, warning=T, echo=FALSE, cache = T}
mixmodel_under_pred_prob <- predict(mixmodel_under_caret_model,
                                as.matrix(mixmodel_dtest_under_20 %>% select(-Outcome)), type="prob")
#plot((roc(response=mixmodel_dtest_20$Outcome, predictor=mixmodel_pred$DEATH)),print.thres="best", print.thres.best.method="closest.topleft")
roc(response=mixmodel_dtest_under_20$Outcome, predictor=mixmodel_under_pred_prob$DEATH, levels = c("ALIVE", "DEATH"), direction = c("<"))
```

##### Predictivy values
```{r undersampled predictive values, echo=FALSE, cache = T}
mixmodel_under_pred <- predict(mixmodel_under_caret_model,
                           as.matrix(mixmodel_dtest_under_20 %>% select(-Outcome)))
confusionMatrix(data = mixmodel_under_pred,
                reference = mixmodel_dtest_under_20$Outcome,
                positive = "DEATH")
```

Confidence intervals of Sensitivity and Specificity (Note that Outcome + means 'DEATH')

```{r undersampled predictive values w CIs, echo=FALSE, cache = T}
library(epiR)
options(digits = 3)
confusionMatrix(data = factor(mixmodel_under_pred,levels=c("DEATH", "ALIVE")) , reference = factor(mixmodel_dtest_under_20$Outcome, levels=c("DEATH", "ALIVE")), positive = "DEATH", mode=c("sens_spec"))$table -> mixmodel_under_matrix_table
epiR::epi.tests(mixmodel_under_matrix_table) -> mixmodel_under_epitests

Sensitivity <- c(mixmodel_under_epitests$elements$se, mixmodel_under_epitests$elements$se.low, mixmodel_under_epitests$elements$se.up)
Specificity <- c(mixmodel_under_epitests$elements$sp, mixmodel_under_epitests$elements$sp.low, mixmodel_under_epitests$elements$sp.up)
`Positive Predictive Value` <- c(mixmodel_under_epitests$elements$ppv, mixmodel_under_epitests$elements$ppv.low, mixmodel_under_epitests$elements$ppv.up)
`Negative Predictive Value` <- c(mixmodel_under_epitests$elements$npv, mixmodel_under_epitests$elements$npv.low, mixmodel_under_epitests$elements$npv.up)
mixmodel_epitests_df <- as.data.frame(rbind(Sensitivity, Specificity, `Positive Predictive Value`, `Negative Predictive Value`))
colnames(mixmodel_epitests_df) <- c("Estimate", "Lower Interval", "Upper Interval")
knitr::kable(mixmodel_epitests_df)
```

#### Model results using all data 

##### AUC

```{r undersampled auc full data, warning=T, echo=FALSE, cache = T}
mixmodel_under_full_data_pred_prob <- predict(mixmodel_under_caret_model,
                                             as.matrix(Farvadin_2_full_data %>% select(-Outcome)), type="prob")
#plot((roc(response=mixmodel_dtest_20$Outcome, predictor=mixmodel_pred$DEATH)),print.thres="best", print.thres.best.method="closest.topleft")
roc(response=Farvadin_2_full_data$Outcome, predictor=mixmodel_under_full_data_pred_prob$DEATH, levels = c("ALIVE", "DEATH"), direction = c("<"))
```

##### Predictivy values
```{r undersampled predictive values full data, echo=FALSE, cache = T}
mixmodel_under_full_data_pred <- predict(mixmodel_under_caret_model,
                                        as.matrix(Farvadin_2_full_data %>% select(-Outcome)))
confusionMatrix(data = mixmodel_under_full_data_pred,
                reference = Farvadin_2_full_data$Outcome,
                positive = "DEATH")
```

Confidence intervals of Sensitivity and Specificity (Note that Outcome + means 'DEATH')

```{r undersampled predictive values w CIs full data, echo=FALSE, cache = T}
library(epiR)
options(digits = 3)
confusionMatrix(data = factor(mixmodel_under_full_data_pred,levels=c("DEATH", "ALIVE")) ,
                reference = factor(Farvadin_2_full_data$Outcome, levels=c("DEATH", "ALIVE")),
                positive = "DEATH", mode=c("sens_spec"))$table -> mixmodel_under_full_data_matrix_table
epiR::epi.tests(mixmodel_under_full_data_matrix_table) -> mixmodel_under_full_data_epitests

Sensitivity <- c(mixmodel_under_full_data_epitests$elements$se, mixmodel_under_full_data_epitests$elements$se.low, mixmodel_under_full_data_epitests$elements$se.up)

Specificity <- c(mixmodel_under_full_data_epitests$elements$sp, mixmodel_under_full_data_epitests$elements$sp.low, mixmodel_under_full_data_epitests$elements$sp.up)

`Positive Predictive Value` <- c(mixmodel_under_full_data_epitests$elements$ppv, mixmodel_under_full_data_epitests$elements$ppv.low, mixmodel_under_full_data_epitests$elements$ppv.up)

`Negative Predictive Value` <- c(mixmodel_under_full_data_epitests$elements$npv, mixmodel_under_full_data_epitests$elements$npv.low, mixmodel_under_full_data_epitests$elements$npv.up)

mixmodel_epitests_full_data_df <- as.data.frame(rbind(Sensitivity, Specificity, `Positive Predictive Value`, `Negative Predictive Value`))
colnames(mixmodel_epitests_full_data_df) <- c("Estimate", "Lower Interval", "Upper Interval")

knitr::kable(mixmodel_epitests_full_data_df)
```


### Oversample ROSE model

Train and test datasets created previoysly using `ROSE` from `ROSE` package and the `createDataPartition(as.factor(Farvadin_2$Outcome), p=0.80, list=FALSE)` also from the `caret` package.

Train data (80%)

```{r ROSE train data loading, echo=FALSE, cache=T}
dtrain_ROSE_80 <- readRDS("Farvadin_2_balanced_ROSE_train_80.Rda")  %>% 
  mutate(Outcome = recode(Outcome, "0" = "ALIVE", "1" = "DEATH"))
summary(dtrain_ROSE_80$Outcome)
```

Test data (20%)

```{r ROSE test data loading, echo=FALSE, warning = F, cache=T}
dtest_ROSE_20 <- readRDS("Farvadin_2_balanced_ROSE_test_20.Rda") %>% 
  mutate(Outcome = recode(Outcome, "0" = "ALIVE", "1" = "DEATH"))
summary(dtest_ROSE_20$Outcome)
```

#### Loading data

```{r ROSE load  data, echo=FALSE, cache=T}
dtrain_ROSE_80 -> mixmodel_dtrain_ROSE_80
dtest_ROSE_20 -> mixmodel_dtest_ROSE_20
```

#### Model trains

```{r ROSE model train, echo=FALSE, cache=T}
# https://github.com/rstudio/rstudio/issues/6692
library(doParallel)
no_cores <- detectCores() - 1
# create the cluster for caret to use
#cl <- makePSOCKcluster(no_cores)
cl <- parallel::makeCluster(no_cores, setup_strategy = "sequential")
registerDoParallel(cl)

registerDoParallel(cl)
mixmodel_ROSE_train_control = trainControl(
  method = "cv",
  number = 5,
  verboseIter = FALSE,
  returnData = FALSE,
  returnResamp = "all", # save losses across all models
  classProbs = TRUE, # set to TRUE for AUC to be computed
  summaryFunction = twoClassSummary,
  allowParallel = TRUE
)

mixmodel_ROSE_tune_grid = expand.grid(
  nrounds = 5000,
  max_depth = c(2, 4, 6, 8, 10),
  eta=c(0.5, 0.3, 0.1, 0.07),
  gamma = c(0, 0.01),
  colsample_bytree=c(1, 0.5),
  min_child_weight=1,
  subsample=c(1, 0.5)
)

# Recode https://stackoverflow.com/a/44131491/9664743
set.seed(0)
mixmodel_ROSE_caret_model = train(
  x = as.matrix(mixmodel_dtrain_ROSE_80 %>% 
                   mutate_at(vars(matches(c("Cat","Respiratorydistress1"))), 
                             function(x) as.numeric(as.character(x))) 
                 %>% select(-Outcome)),
  y = mixmodel_dtrain_ROSE_80$Outcome,
  trControl = mixmodel_ROSE_train_control,
  tuneGrid = mixmodel_ROSE_tune_grid,
  metric= "ROC",
  method = "xgbTree"
)
registerDoSEQ()
mixmodel_ROSE_caret_model$bestTune
```

```{r ROSE care model save}
saveRDS(mixmodel_ROSE_caret_model, file="mixmodel_ROSE_caret_train_20121204.model")
```
##### Variables importance

```{r ROSE variable importance, echo=FALSE, cache=T}
mixmodel_ROSE_var_importance <- varImp(mixmodel_ROSE_caret_model, scale = FALSE, dig=3)
mixmodel_ROSE_var_importance
```


```{r ROSE variable importance plot, echo=FALSE, cache=T}

## PT                   0.36791
## Neutrophilspercent   0.11763
## BloodUreaNitrogen    0.09194
## APTT                 0.08906
## ASTALTRatio          0.07562
## CRP                  0.06195
## Age                  0.05177
## lymphocytespercent   0.04336
## Respiratorydistress1 0.04020
## CatHaematocrit_0     0.03106
## CatHaematocrit_1     0.02951
## CatHaematocrit_2     0.00000

rownames(mixmodel_ROSE_var_importance$importance) <- 
  c("PT",
    "Neutrophils % ",
    "Blood Urea Nitrogen",
    "APTT",
    "AST/ALT Ratio",
    "CRP",
    "Age",
    "Lymphocytes %",
    "Respiratory distress",
    "Haematocrit - Less",
    "Haematocrit - Normal",
    "Haematocrit - More")
# plot(mixmodel_var_importance)
ggplot(mixmodel_ROSE_var_importance)
```

#### Model results using oversample TEST set

##### AUC

```{r ROSE auc, warning=T, echo=FALSE, cache=T}
mixmodel_ROSE_pred_prob <- predict(mixmodel_ROSE_caret_model,
                                as.matrix(mixmodel_dtest_ROSE_20 %>% 
                                            mutate_at(vars(matches(c("Cat","Respiratorydistress1"))), 
                                                      function(x) as.numeric(as.character(x))) 
                                          %>% select(-Outcome)), type="prob")
#plot((roc(response=mixmodel_dtest_20$Outcome, predictor=mixmodel_pred$DEATH)),print.thres="best", print.thres.best.method="closest.topleft")
roc(response=mixmodel_dtest_ROSE_20$Outcome, predictor=mixmodel_ROSE_pred_prob$DEATH, levels = c("ALIVE", "DEATH"), direction = c("<"))
```

##### Predictivy values
```{r ROSE predictive values, echo=FALSE, cache=T}
mixmodel_ROSE_pred <- predict(mixmodel_ROSE_caret_model,
                           as.matrix(mixmodel_dtest_ROSE_20 %>% 
                                            mutate_at(vars(matches(c("Cat","Respiratorydistress1"))), 
                                                      function(x) as.numeric(as.character(x))) 
                                          %>% select(-Outcome)))
confusionMatrix(data = mixmodel_ROSE_pred,
                reference = mixmodel_dtest_ROSE_20$Outcome,
                positive = "DEATH")
```

Confidence intervals of Sensitivity and Specificity (Note that Outcome + means 'DEATH')

```{r ROSE predictive values w CIs, echo=FALSE, cache=T}
library(epiR)
options(digits = 3)
confusionMatrix(data = factor(mixmodel_ROSE_pred,levels=c("DEATH", "ALIVE")) , reference = factor(mixmodel_dtest_ROSE_20$Outcome, levels=c("DEATH", "ALIVE")), positive = "DEATH", mode=c("sens_spec"))$table -> mixmodel_ROSE_matrix_table
epiR::epi.tests(mixmodel_ROSE_matrix_table) -> mixmodel_ROSE_epitests

Sensitivity <- c(mixmodel_ROSE_epitests$elements$se, mixmodel_ROSE_epitests$elements$se.low, mixmodel_ROSE_epitests$elements$se.up)
Specificity <- c(mixmodel_ROSE_epitests$elements$sp, mixmodel_ROSE_epitests$elements$sp.low, mixmodel_ROSE_epitests$elements$sp.up)
`Positive Predictive Value` <- c(mixmodel_ROSE_epitests$elements$ppv, mixmodel_ROSE_epitests$elements$ppv.low, mixmodel_ROSE_epitests$elements$ppv.up)
`Negative Predictive Value` <- c(mixmodel_ROSE_epitests$elements$npv, mixmodel_ROSE_epitests$elements$npv.low, mixmodel_ROSE_epitests$elements$npv.up)
mixmodel_epitests_df <- as.data.frame(rbind(Sensitivity, Specificity, `Positive Predictive Value`, `Negative Predictive Value`))
colnames(mixmodel_epitests_df) <- c("Estimate", "Lower Interval", "Upper Interval")
knitr::kable(mixmodel_epitests_df)
```

#### Model results using all data 

##### AUC

```{r ROSE auc full data, warning=T, echo=FALSE, cache=T}
mixmodel_ROSE_full_data_pred_prob <- predict(mixmodel_ROSE_caret_model,
                                             as.matrix(Farvadin_2_full_data %>% select(-Outcome)), type="prob")
#plot((roc(response=mixmodel_dtest_20$Outcome, predictor=mixmodel_pred$DEATH)),print.thres="best", print.thres.best.method="closest.topleft")
roc(response=Farvadin_2_full_data$Outcome, predictor=mixmodel_ROSE_full_data_pred_prob$DEATH, levels = c("ALIVE", "DEATH"), direction = c("<"))
```

##### Predictivy values
```{r ROSE predictive values full data, echo=FALSE, cache=T}
mixmodel_ROSE_full_data_pred <- predict(mixmodel_ROSE_caret_model,
                                        as.matrix(Farvadin_2_full_data %>% select(-Outcome)))
confusionMatrix(data = mixmodel_ROSE_full_data_pred,
                reference = Farvadin_2_full_data$Outcome,
                positive = "DEATH")
```

Confidence intervals of Sensitivity and Specificity (Note that Outcome + means 'DEATH')

```{r ROSE predictive values w CIs full data, echo=FALSE, cache=T}
library(epiR)
options(digits = 3)
confusionMatrix(data = factor(mixmodel_ROSE_full_data_pred,levels=c("DEATH", "ALIVE")) ,
                reference = factor(Farvadin_2_full_data$Outcome, levels=c("DEATH", "ALIVE")),
                positive = "DEATH", mode=c("sens_spec"))$table -> mixmodel_ROSE_full_data_matrix_table
epiR::epi.tests(mixmodel_ROSE_full_data_matrix_table) -> mixmodel_ROSE_full_data_epitests

Sensitivity <- c(mixmodel_ROSE_full_data_epitests$elements$se, mixmodel_ROSE_full_data_epitests$elements$se.low, mixmodel_ROSE_full_data_epitests$elements$se.up)

Specificity <- c(mixmodel_ROSE_full_data_epitests$elements$sp, mixmodel_ROSE_full_data_epitests$elements$sp.low, mixmodel_ROSE_full_data_epitests$elements$sp.up)

`Positive Predictive Value` <- c(mixmodel_ROSE_full_data_epitests$elements$ppv, mixmodel_ROSE_full_data_epitests$elements$ppv.low, mixmodel_ROSE_full_data_epitests$elements$ppv.up)

`Negative Predictive Value` <- c(mixmodel_ROSE_full_data_epitests$elements$npv, mixmodel_ROSE_full_data_epitests$elements$npv.low, mixmodel_ROSE_full_data_epitests$elements$npv.up)

mixmodel_epitests_full_data_df <- as.data.frame(rbind(Sensitivity, Specificity, `Positive Predictive Value`, `Negative Predictive Value`))
colnames(mixmodel_epitests_full_data_df) <- c("Estimate", "Lower Interval", "Upper Interval")

knitr::kable(mixmodel_epitests_full_data_df)
```
