---
title: "COVID-19 Analyses"
output: html_document
---

```{r setup, include=FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(haven)

library(caret)
library(pROC)
```

##  XGBoost exploration {.tabset}

Train and test datasets created previoysly using `reateDataPartition(as.factor(Farvadin_2$Outcome), p=0.80, list=FALSE)`

Train data (80%)

```{r train data loading, cache = TRUE}
dtrain_80 <- readRDS("Farvadin_2_train_80.Rda")  %>% 
  mutate(Outcome = recode(Outcome, "0" = "ALIVE", "1" = "DEATH"))
summary(dtrain_80$Outcome)
```

Test data (20%)

```{r test data loading, warning = F, cache = TRUE}
dtest_20 <- readRDS("Farvadin_2_test_20.Rda") %>% 
  mutate(Outcome = recode(Outcome, "0" = "ALIVE", "1" = "DEATH"))
summary(dtest_20$Outcome)
```

### Best Total model

Independet variables:

* `PT`
* `Neutrophilspercent`
* `APTT`
* `CatHaematocrit`
* `Platelets`
* `BloodUreaNitrogen`
* `Creatinine`
* `Age`
* `ChronicKidneyDisease1`
* `ASTALTRatio`
* `HB`
* `TotalPrimarySigns`

#### Loading data

```{r best total load  data, cache = TRUE}
dtrain_80 %>%
  select(Outcome,
         PT,
         Neutrophilspercent,
         APTT,
         CatHaematocrit,
         Platelets,
         BloodUreaNitrogen,
         Creatinine,
         Age,
         ChronicKidneyDisease1,
         ASTALTRatio,
         HB,
         TotalPrimarySigns) %>% 
  # filter(complete.cases(.)) %>% 
  fastDummies::dummy_columns(select_columns = c("CatHaematocrit"), 
                             ignore_na = T,
                             remove_selected_columns = T) -> best_total_dtrain_80

dtest_20 %>%
  select(Outcome,
         PT,
         Neutrophilspercent,
         APTT,
         CatHaematocrit,
         Platelets,
         BloodUreaNitrogen,
         Creatinine,
         Age,
         ChronicKidneyDisease1,
         ASTALTRatio,
         HB,
         TotalPrimarySigns) %>% 
  # filter(complete.cases(.)) %>% 
  fastDummies::dummy_columns(select_columns = c("CatHaematocrit"), 
                             ignore_na = T,
                             remove_selected_columns = T) -> best_total_dtest_20

```

#### Model train

```{r best_total model train, cache = TRUE}
library(doParallel)
cl <- makePSOCKcluster(detectCores())
registerDoParallel(cl)

best_total_train_control = trainControl(
  method = "cv",
  number = 5,
  verboseIter = FALSE,
  returnData = FALSE,
  returnResamp = "all", # save losses across all models
  classProbs = TRUE, # set to TRUE for AUC to be computed
  summaryFunction = twoClassSummary,
  allowParallel = TRUE
)

best_total_tune_grid = expand.grid(
  nrounds = 5000,
  max_depth = c(2, 4, 6, 8, 10),
  eta=c(0.5, 0.3, 0.1, 0.07),
  gamma = c(0, 0.01),
  colsample_bytree=c(1, 0.5),
  min_child_weight=1,
  subsample=c(1, 0.5)
)

# Recode https://stackoverflow.com/a/44131491/9664743
set.seed(0)
best_total_caret_model = train(
  x = as.matrix(best_total_dtrain_80 %>% select(-Outcome)),
  y = best_total_dtrain_80$Outcome,
  trControl = best_total_train_control,
  tuneGrid = best_total_tune_grid,
  metric= "ROC",
  method = "xgbTree"
)
registerDoSEQ()

best_total_caret_model$bestTune
```

```{r best total caret model save}
saveRDS(best_total_caret_model, file="best_total_caret_train.model")
```

#### Model results

##### AUC

```{r best total auc, warning=T, echo=FALSE, cache = T}
best_total_pred_prob <- predict(best_total_caret_model,
                                as.matrix(best_total_dtest_20 %>% select(-Outcome)), type="prob")
#plot((roc(response=best_total_dtest_20$Outcome, predictor=best_total_pred$DEATH)),print.thres="best", print.thres.best.method="closest.topleft")
roc(response=best_total_dtest_20$Outcome, predictor=best_total_pred_prob$DEATH, levels = c("ALIVE", "DEATH"), direction = c("<"))
```

##### Predictivy values
```{r best total predictive values, cache = T}
best_total_pred <- predict(best_total_caret_model,
                           as.matrix(best_total_dtest_20 %>% select(-Outcome)))
confusionMatrix(data = best_total_pred,
                reference = best_total_dtest_20$Outcome,
                positive = "DEATH")
```

Confidence intervals of Sensitivity and Specificity (Note that Outcome + means 'DEATH')

```{r best_total predictive values w CIs, cache = T}
confusionMatrix(data = factor(best_total_pred,levels=c("DEATH", "ALIVE")) , reference = factor(best_total_dtest_20$Outcome, levels=c("DEATH", "ALIVE")), positive = "DEATH", mode=c("sens_spec"))$table -> best_total_matrix_table
epiR::epi.tests(best_total_matrix_table) -> best_total_epitests

Sensitivity <- c(best_total_epitests$elements$se, best_total_epitests$elements$se.low, best_total_epitests$elements$se.up)
Specificity <- c(best_total_epitests$elements$sp, best_total_epitests$elements$sp.low, best_total_epitests$elements$sp.up)
`Positive Predictive Value` <- c(best_total_epitests$elements$ppv, best_total_epitests$elements$ppv.low, best_total_epitests$elements$ppv.up)
`Negative Predictive Value` <- c(best_total_epitests$elements$npv, best_total_epitests$elements$npv.low, best_total_epitests$elements$npv.up)
best_total_epitests_df <- as.data.frame(rbind(Sensitivity, Specificity, `Positive Predictive Value`, `Negative Predictive Value`))
colnames(best_total_epitests_df) <- c("Estimate", "Lower Interval", "Upper Interval")
knitr::kable(best_total_epitests_df)
```

##### Variables importance

```{r best total variable importance, echo=FALSE, cache=T}
best_total_var_importance <- varImp(best_total_caret_model, scale = FALSE)
plot(best_total_var_importance)
```

### Clinical model

Independet variables:

* `Age`
* `PrimarySignFever1`
* `PrimarySignMuscularpain`
* `PrimarySignCough1`
* `PrimarySignDecreasedconsciousnes`
* `Respiratorydistress1`
* `Cancer1`
* `Cardiovasculardisease1`
* `ChronicKidneyDisease1`
* `Diabetes1`
* `OtherLungDisease1`
* `OtherChronicDisease1`

#### Loading data

```{r clinical load  data, cache = TRUE}
dtrain_80 %>%
  select(Outcome,
         Age,
         PrimarySignFever1,
         PrimarySignMuscularpain,
         PrimarySignCough1,
         PrimarySignDecreasedconsciousnes,
         Respiratorydistress1,
         Cancer1,
         Cardiovasculardisease1,
         ChronicKidneyDisease1,
         Diabetes1,
         OtherLungDisease1,
         OtherChronicDisease1) -> clinical_dtrain_80

dtest_20 %>%
  select(Outcome,
         Age,
         PrimarySignFever1,
         PrimarySignMuscularpain,
         PrimarySignCough1,
         PrimarySignDecreasedconsciousnes,
         Respiratorydistress1,
         Cancer1,
         Cardiovasculardisease1,
         ChronicKidneyDisease1,
         Diabetes1,
         OtherLungDisease1,
         OtherChronicDisease1) -> clinical_dtest_20

```

#### Model train

```{r clinical model train, cache = TRUE}
library(doParallel)
cl <- makePSOCKcluster(detectCores())
registerDoParallel(cl)
clinical_train_control = trainControl(
  method = "cv",
  number = 5,
  verboseIter = FALSE,
  returnData = FALSE,
  returnResamp = "all", # save losses across all models
  classProbs = TRUE, # set to TRUE for AUC to be computed
  summaryFunction = twoClassSummary,
  allowParallel = TRUE
)

clinical_tune_grid = expand.grid(
  nrounds = 5000,
  max_depth = c(2, 4, 6, 8, 10),
  eta=c(0.5, 0.3, 0.1, 0.07),
  gamma = c(0, 0.01),
  colsample_bytree=c(1, 0.5),
  min_child_weight=1,
  subsample=c(1, 0.5)
)

# Recode https://stackoverflow.com/a/44131491/9664743
set.seed(0)
clinical_caret_model = train(
  x = as.matrix(clinical_dtrain_80 %>% select(-Outcome)),
  y = clinical_dtrain_80$Outcome,
  trControl = clinical_train_control,
  tuneGrid = clinical_tune_grid,
  metric= "ROC",
  method = "xgbTree"
)
registerDoSEQ()
clinical_caret_model$bestTune
```

```{r clinical care model save}
saveRDS(clinical_caret_model, file="clinical_caret_train.model")
```


#### Model results

##### AUC

```{r clinical auc, warning=T, echo=FALSE, cache = T}
clinical_pred_prob <- predict(clinical_caret_model,
                                as.matrix(clinical_dtest_20 %>% select(-Outcome)), type="prob")
#plot((roc(response=clinical_dtest_20$Outcome, predictor=clinical_pred$DEATH)),print.thres="best", print.thres.best.method="closest.topleft")
roc(response=clinical_dtest_20$Outcome, predictor=clinical_pred_prob$DEATH, levels = c("ALIVE", "DEATH"), direction = c("<"))
```

##### Predictivy values
```{r clinical predictive values, cache = T}
clinical_pred <- predict(clinical_caret_model,
                           as.matrix(clinical_dtest_20 %>% select(-Outcome)))
confusionMatrix(data = clinical_pred,
                reference = clinical_dtest_20$Outcome,
                positive = "DEATH")
```

Confidence intervals of Sensitivity and Specificity (Note that Outcome + means 'DEATH')

```{r clinical predictive values w CIs, cache = T}
library(epiR)
options(digits = 3)
confusionMatrix(data = factor(clinical_pred,levels=c("DEATH", "ALIVE")) , reference = factor(clinical_dtest_20$Outcome, levels=c("DEATH", "ALIVE")), positive = "DEATH", mode=c("sens_spec"))$table -> clinical_matrix_table
epiR::epi.tests(clinical_matrix_table) -> clinical_epitests

Sensitivity <- c(clinical_epitests$elements$se, clinical_epitests$elements$se.low, clinical_epitests$elements$se.up)
Specificity <- c(clinical_epitests$elements$sp, clinical_epitests$elements$sp.low, clinical_epitests$elements$sp.up)
`Positive Predictive Value` <- c(clinical_epitests$elements$ppv, clinical_epitests$elements$ppv.low, clinical_epitests$elements$ppv.up)
`Negative Predictive Value` <- c(clinical_epitests$elements$npv, clinical_epitests$elements$npv.low, clinical_epitests$elements$npv.up)
clinical_epitests_df <- as.data.frame(rbind(Sensitivity, Specificity, `Positive Predictive Value`, `Negative Predictive Value`))
colnames(clinical_epitests_df) <- c("Estimate", "Lower Interval", "Upper Interval")
knitr::kable(clinical_epitests_df)
```

##### Variables importance

```{r clinical variable importance, echo=FALSE, cache=T}
clinical_var_importance <- varImp(clinical_caret_model, scale = FALSE)
plot(clinical_var_importance)
```

### Third model

Independet variables:

* `PT`
* `Neutrophilspercent`
* `ASTALTRatio`
* `APTT`
* `CatHaematocrit`
* `Platelets`
* `BloodUreaNitrogen`
* `Creatinine`
* `Age`
* `Cancer1`
* `INR`
* `CatHB`
* `TotalPrimarySigns`
* `CRP`


#### Loading data

```{r third_model load  data, cache = TRUE}
dtrain_80 %>%
  select(Outcome,
         PT,
         Neutrophilspercent,
         ASTALTRatio,
         APTT,
         CatHaematocrit,
         Platelets,
         BloodUreaNitrogen,
         Creatinine,
         Age,
         Cancer1,
         INR,
         CatHB,
         TotalPrimarySigns,
         CRP) %>% 
  # filter(complete.cases(.)) %>% 
  fastDummies::dummy_columns(select_columns = c("CatHaematocrit", "CatHB"), 
                             ignore_na = T,
                             remove_selected_columns = T) -> third_model_dtrain_80

dtest_20 %>%
  select(Outcome,
         PT,
         Neutrophilspercent,
         ASTALTRatio,
         APTT,
         CatHaematocrit,
         Platelets,
         BloodUreaNitrogen,
         Creatinine,
         Age,
         Cancer1,
         INR,
         CatHB,
         TotalPrimarySigns,
         CRP) %>% 
  # filter(complete.cases(.)) %>% 
  fastDummies::dummy_columns(select_columns = c("CatHaematocrit", "CatHB"), 
                             ignore_na = T,
                             remove_selected_columns = T) -> third_model_dtest_20

```

#### Model train

```{r third_model model train, cache = TRUE}
library(doParallel)
cl <- makePSOCKcluster(detectCores())
registerDoParallel(cl)
third_model_train_control = trainControl(
  method = "cv",
  number = 5,
  verboseIter = FALSE,
  returnData = FALSE,
  returnResamp = "all", # save losses across all models
  classProbs = TRUE, # set to TRUE for AUC to be computed
  summaryFunction = twoClassSummary,
  allowParallel = TRUE
)

third_model_tune_grid = expand.grid(
  nrounds = 5000,
  max_depth = c(2, 4, 6, 8, 10),
  eta=c(0.5, 0.3, 0.1, 0.07),
  gamma = c(0, 0.01),
  colsample_bytree=c(1, 0.5),
  min_child_weight=1,
  subsample=c(1, 0.5)
)

# Recode https://stackoverflow.com/a/44131491/9664743
set.seed(0)
third_model_caret_model = train(
  x = as.matrix(third_model_dtrain_80 %>% select(-Outcome)),
  y = third_model_dtrain_80$Outcome,
  trControl = third_model_train_control,
  tuneGrid = third_model_tune_grid,
  metric= "ROC",
  method = "xgbTree"
)
registerDoSEQ()
third_model_caret_model$bestTune
```

```{r third_model care model save}
saveRDS(third_model_caret_model, file="third_model_caret_train.model")
```


#### Model results

##### AUC

```{r third_model auc, warning=T, echo=FALSE, cache = T}
third_model_pred_prob <- predict(third_model_caret_model,
                                as.matrix(third_model_dtest_20 %>% select(-Outcome)), type="prob")
#plot((roc(response=third_model_dtest_20$Outcome, predictor=third_model_pred$DEATH)),print.thres="best", print.thres.best.method="closest.topleft")
roc(response=third_model_dtest_20$Outcome, predictor=third_model_pred_prob$DEATH, levels = c("ALIVE", "DEATH"), direction = c("<"))
```

##### Predictivy values
```{r third_model predictive values, cache = T}
third_model_pred <- predict(third_model_caret_model,
                           as.matrix(third_model_dtest_20 %>% select(-Outcome)))
confusionMatrix(data = third_model_pred,
                reference = third_model_dtest_20$Outcome,
                positive = "DEATH")
```

Confidence intervals of Sensitivity and Specificity (Note that Outcome + means 'DEATH')

```{r third_model predictive values w CIs, cache = T}
library(epiR)
options(digits = 3)
confusionMatrix(data = factor(third_model_pred,levels=c("DEATH", "ALIVE")) , reference = factor(third_model_dtest_20$Outcome, levels=c("DEATH", "ALIVE")), positive = "DEATH", mode=c("sens_spec"))$table -> third_model_matrix_table
epiR::epi.tests(third_model_matrix_table) -> third_model_epitests

Sensitivity <- c(third_model_epitests$elements$se, third_model_epitests$elements$se.low, third_model_epitests$elements$se.up)
Specificity <- c(third_model_epitests$elements$sp, third_model_epitests$elements$sp.low, third_model_epitests$elements$sp.up)
`Positive Predictive Value` <- c(third_model_epitests$elements$ppv, third_model_epitests$elements$ppv.low, third_model_epitests$elements$ppv.up)
`Negative Predictive Value` <- c(third_model_epitests$elements$npv, third_model_epitests$elements$npv.low, third_model_epitests$elements$npv.up)
third_model_epitests_df <- as.data.frame(rbind(Sensitivity, Specificity, `Positive Predictive Value`, `Negative Predictive Value`))
colnames(third_model_epitests_df) <- c("Estimate", "Lower Interval", "Upper Interval")
knitr::kable(third_model_epitests_df)
```

##### Variables importance

```{r third_model variable importance, echo=FALSE, cache=T}
third_model_var_importance <- varImp(third_model_caret_model, scale = FALSE)
plot(third_model_var_importance)
```

### Total Stepwise model

Independet variables:

* `PT`
* `Neutrophilspercent`
* `Res`
* `CatHaematocrit`
* `Age`
* `CRP`
* `Respiratorydistress1`
* `ASTALTRatio`
* `APTT`
* `BloodUreaNitrogen`


#### Loading data

```{r total_stepwise load  data, cache = TRUE}
dtrain_80 %>%
  select(Outcome,
         PT,
         Neutrophilspercent,
         Respiratorydistress1,
         Cardiovasculardisease1,
         Haematocrit,
         Age,
         CRP) -> total_stepwise_dtrain_80

dtest_20 %>%
  select(Outcome,
         PT,
         Neutrophilspercent,
         Respiratorydistress1,
         Cardiovasculardisease1,
         Haematocrit,
         Age,
         CRP) -> total_stepwise_dtest_20

```

#### Caret tuning

```{r total_stepwise model train, cache = TRUE}
library(doParallel)
cl <- makePSOCKcluster(detectCores())
registerDoParallel(cl)
total_stepwise_train_control = trainControl(
  method = "cv",
  number = 5,
  verboseIter = FALSE,
  returnData = FALSE,
  returnResamp = "all", # save losses across all models
  classProbs = TRUE, # set to TRUE for AUC to be computed
  summaryFunction = twoClassSummary,
  allowParallel = TRUE
)

total_stepwise_tune_grid = expand.grid(
  nrounds = 5000,
  max_depth = c(2, 4, 6, 8, 10),
  eta=c(0.5, 0.3, 0.1, 0.07),
  gamma = c(0, 0.01),
  colsample_bytree=c(1, 0.5),
  min_child_weight=1,
  subsample=c(1, 0.5)
)

# Recode https://stackoverflow.com/a/44131491/9664743
set.seed(0)
total_stepwise_caret_model = train(
  x = as.matrix(total_stepwise_dtrain_80 %>% select(-Outcome)),
  y = total_stepwise_dtrain_80$Outcome,
  trControl = total_stepwise_train_control,
  tuneGrid = total_stepwise_tune_grid,
  metric= "ROC",
  method = "xgbTree"
)
registerDoSEQ()
total_stepwise_caret_model$bestTune
```

```{r total_stepwise care model save}
saveRDS(total_stepwise_caret_model, file="total_stepwise_caret_train.model")
```


#### Model results

##### AUC

```{r total_stepwise auc, warning=T, echo=FALSE, cache = T}
total_stepwise_pred_prob <- predict(total_stepwise_caret_model,
                                as.matrix(total_stepwise_dtest_20 %>% select(-Outcome)), type="prob")
#plot((roc(response=total_stepwise_dtest_20$Outcome, predictor=total_stepwise_pred$DEATH)),print.thres="best", print.thres.best.method="closest.topleft")
roc(response=total_stepwise_dtest_20$Outcome, predictor=total_stepwise_pred_prob$DEATH, levels = c("ALIVE", "DEATH"), direction = c("<"))
```

##### Predictivy values
```{r total_stepwise predictive values, cache = T}
total_stepwise_pred <- predict(total_stepwise_caret_model,
                           as.matrix(total_stepwise_dtest_20 %>% select(-Outcome)))
confusionMatrix(data = total_stepwise_pred,
                reference = total_stepwise_dtest_20$Outcome,
                positive = "DEATH")
```

Confidence intervals of Sensitivity and Specificity (Note that Outcome + means 'DEATH')

```{r total_stepwise predictive values w CIs, cache = T}
library(epiR)
options(digits = 3)
confusionMatrix(data = factor(total_stepwise_pred,levels=c("DEATH", "ALIVE")) , reference = factor(total_stepwise_dtest_20$Outcome, levels=c("DEATH", "ALIVE")), positive = "DEATH", mode=c("sens_spec"))$table -> total_stepwise_matrix_table
epiR::epi.tests(total_stepwise_matrix_table) -> total_stepwise_epitests

Sensitivity <- c(total_stepwise_epitests$elements$se, total_stepwise_epitests$elements$se.low, total_stepwise_epitests$elements$se.up)
Specificity <- c(total_stepwise_epitests$elements$sp, total_stepwise_epitests$elements$sp.low, total_stepwise_epitests$elements$sp.up)
`Positive Predictive Value` <- c(total_stepwise_epitests$elements$ppv, total_stepwise_epitests$elements$ppv.low, total_stepwise_epitests$elements$ppv.up)
`Negative Predictive Value` <- c(total_stepwise_epitests$elements$npv, total_stepwise_epitests$elements$npv.low, total_stepwise_epitests$elements$npv.up)
total_stepwise_epitests_df <- as.data.frame(rbind(Sensitivity, Specificity, `Positive Predictive Value`, `Negative Predictive Value`))
colnames(total_stepwise_epitests_df) <- c("Estimate", "Lower Interval", "Upper Interval")
knitr::kable(total_stepwise_epitests_df)
```

##### Variables importance

```{r total_stepwise variable importance, echo=FALSE, cache=T}
total_stepwise_var_importance <- varImp(total_stepwise_caret_model, scale = FALSE)
plot(total_stepwise_var_importance)
```

### Mixmodel model

Independet variables:

* `PT`
* `Neutrophilspercent`
* `lymphocytespercent`
* `CatHaematocrit`
* `Age`
* `CRP`
* `Respiratorydistress1`
* `ASTALTRatio`
* `APTT`
* `BloodUreaNitrogen`


#### Loading data

```{r mixmodel load  data, cache = TRUE}
dtrain_80 %>%
  select(Outcome,
         PT,
         Neutrophilspercent,
         lymphocytespercent,
         CatHaematocrit,
         Age,
         CRP,
         Respiratorydistress1,
         ASTALTRatio,
         APTT,
         BloodUreaNitrogen) %>% 
  # filter(complete.cases(.)) %>% 
  fastDummies::dummy_columns(select_columns = c("CatHaematocrit", "CatHB"), 
                             ignore_na = T,
                             remove_selected_columns = T) -> mixmodel_dtrain_80

dtest_20 %>%
  select(Outcome,
         PT,
         Neutrophilspercent,
         lymphocytespercent,
         CatHaematocrit,
         Age,
         CRP,
         Respiratorydistress1,
         ASTALTRatio,
         APTT,
         BloodUreaNitrogen) %>% 
  # filter(complete.cases(.)) %>% 
  fastDummies::dummy_columns(select_columns = c("CatHaematocrit", "CatHB"), 
                             ignore_na = T,
                             remove_selected_columns = T) -> mixmodel_dtest_20

```

#### Model trains

```{r mixmodel model train, cache = TRUE}
library(doParallel)
cl <- makePSOCKcluster(detectCores())
registerDoParallel(cl)
mixmodel_train_control = trainControl(
  method = "cv",
  number = 5,
  verboseIter = FALSE,
  returnData = FALSE,
  returnResamp = "all", # save losses across all models
  classProbs = TRUE, # set to TRUE for AUC to be computed
  summaryFunction = twoClassSummary,
  allowParallel = TRUE
)

mixmodel_tune_grid = expand.grid(
  nrounds = 5000,
  max_depth = c(2, 4, 6, 8, 10),
  eta=c(0.5, 0.3, 0.1, 0.07),
  gamma = c(0, 0.01),
  colsample_bytree=c(1, 0.5),
  min_child_weight=1,
  subsample=c(1, 0.5)
)

# Recode https://stackoverflow.com/a/44131491/9664743
set.seed(0)
mixmodel_caret_model = train(
  x = as.matrix(mixmodel_dtrain_80 %>% select(-Outcome)),
  y = mixmodel_dtrain_80$Outcome,
  trControl = mixmodel_train_control,
  tuneGrid = mixmodel_tune_grid,
  metric= "ROC",
  method = "xgbTree"
)
registerDoSEQ()
mixmodel_caret_model$bestTune
```

```{r mixmodel care model save}
saveRDS(mixmodel_caret_model, file="mixmodel_caret_train.model")
```


#### Model results

##### AUC

```{r mixmodel auc, warning=T, echo=FALSE, cache = T}
mixmodel_pred_prob <- predict(mixmodel_caret_model,
                                as.matrix(mixmodel_dtest_20 %>% select(-Outcome)), type="prob")
#plot((roc(response=mixmodel_dtest_20$Outcome, predictor=mixmodel_pred$DEATH)),print.thres="best", print.thres.best.method="closest.topleft")
roc(response=mixmodel_dtest_20$Outcome, predictor=mixmodel_pred_prob$DEATH, levels = c("ALIVE", "DEATH"), direction = c("<"))
```

##### Predictivy values
```{r mixmodel predictive values, cache = T}
mixmodel_pred <- predict(mixmodel_caret_model,
                           as.matrix(mixmodel_dtest_20 %>% select(-Outcome)))
confusionMatrix(data = mixmodel_pred,
                reference = mixmodel_dtest_20$Outcome,
                positive = "DEATH")
```

Confidence intervals of Sensitivity and Specificity (Note that Outcome + means 'DEATH')

```{r mixmodel predictive values w CIs, cache = T}
library(epiR)
options(digits = 3)
confusionMatrix(data = factor(mixmodel_pred,levels=c("DEATH", "ALIVE")) , reference = factor(mixmodel_dtest_20$Outcome, levels=c("DEATH", "ALIVE")), positive = "DEATH", mode=c("sens_spec"))$table -> mixmodel_matrix_table
epiR::epi.tests(mixmodel_matrix_table) -> mixmodel_epitests

Sensitivity <- c(mixmodel_epitests$elements$se, mixmodel_epitests$elements$se.low, mixmodel_epitests$elements$se.up)
Specificity <- c(mixmodel_epitests$elements$sp, mixmodel_epitests$elements$sp.low, mixmodel_epitests$elements$sp.up)
`Positive Predictive Value` <- c(mixmodel_epitests$elements$ppv, mixmodel_epitests$elements$ppv.low, mixmodel_epitests$elements$ppv.up)
`Negative Predictive Value` <- c(mixmodel_epitests$elements$npv, mixmodel_epitests$elements$npv.low, mixmodel_epitests$elements$npv.up)
mixmodel_epitests_df <- as.data.frame(rbind(Sensitivity, Specificity, `Positive Predictive Value`, `Negative Predictive Value`))
colnames(mixmodel_epitests_df) <- c("Estimate", "Lower Interval", "Upper Interval")
knitr::kable(mixmodel_epitests_df)
```

##### Variables importance

```{r mixmodel variable importance, echo=FALSE, cache=T}
mixmodel_var_importance <- varImp(mixmodel_caret_model, scale = FALSE, dig=3)
mixmodel_var_importance
```


```{r mixmodel variable importance plot, echo=FALSE, cache=T}

## BloodUreaNitrogen    0.192687
## ASTALTRatio          0.163023
## Neutrophilspercent   0.135212
## APTT                 0.132159
## CRP                  0.111953
## Age                  0.092079
## lymphocytespercent   0.077785
## PT                   0.072139
## Respiratorydistress1 0.009341
## CatHaematocrit_2     0.007466
## CatHaematocrit_1     0.003401
## CatHaematocrit_0     0.002753

rownames(mixmodel_var_importance$importance) <- 
  c("Blood Urea Nitrogen",
    "AST/ALT Ratio",
    "Neutrophils % ",
    "APTT",
    "CRP",
    "Age",
    "Lymphocytes %",
    "PT",
    "Respiratory distress",
    "Haematocrit - More",
    "Haematocrit - Normal",
    "Haematocrit - Less")
# plot(mixmodel_var_importance)
ggplot(mixmodel_var_importance)
```

### Clinical Stepwise model

Independet variables:

* `PT`
* `Neutrophilspercent`
* `ASTALTRatio`
* `APTT`
* `CatHaematocrit`
* `Platelets`
* `BloodUreaNitrogen`
* `Creatinine`
* `Age`
* `Cancer1`
* `INR`
* `CatHB`
* `CRP`
* `TotalPrimarySigns`


#### Loading data

```{r clinical_stepwise load  data, cache = TRUE}
dtrain_80 %>%
  select(Outcome,
         SEX1,
         Totalcommorbidities1,
         Diabetes1,
         PrimarySignCough1,
         ChronicKidneyDisease1,
         Respiratorydistress1) %>% 
  # filter(complete.cases(.)) %>% 
  fastDummies::dummy_columns(select_columns = c("Diabetes1", "PrimarySignCough1", "ChronicKidneyDisease1", "Respiratorydistress1"), 
                             ignore_na = T,
                             remove_selected_columns = T) -> clinical_stepwise_dtrain_80

dtest_20 %>%
  select(Outcome,
         SEX1,
         Totalcommorbidities1,
         Diabetes1,
         PrimarySignCough1,
         ChronicKidneyDisease1,
         Respiratorydistress1) %>% 
  # filter(complete.cases(.)) %>% 
  fastDummies::dummy_columns(select_columns = c("Diabetes1", "PrimarySignCough1", "ChronicKidneyDisease1", "Respiratorydistress1"), 
                             ignore_na = T,
                             remove_selected_columns = T) -> clinical_stepwise_dtest_20

```

#### Model train

```{r clinical_stepwise model train, cache = TRUE}
library(doParallel)
cl <- makePSOCKcluster(detectCores())
registerDoParallel(cl)
clinical_stepwise_train_control = trainControl(
  method = "cv",
  number = 5,
  verboseIter = FALSE,
  returnData = FALSE,
  returnResamp = "all", # save losses across all models
  classProbs = TRUE, # set to TRUE for AUC to be computed
  summaryFunction = twoClassSummary,
  allowParallel = TRUE
)

clinical_stepwise_tune_grid = expand.grid(
  nrounds = 5000,
  max_depth = c(2, 4, 6, 8, 10),
  eta=c(0.5, 0.3, 0.1, 0.07),
  gamma = c(0, 0.01),
  colsample_bytree=c(1, 0.5),
  min_child_weight=1,
  subsample=c(1, 0.5)
)

# Recode https://stackoverflow.com/a/44131491/9664743
set.seed(0)
clinical_stepwise_caret_model = train(
  x = as.matrix(clinical_stepwise_dtrain_80 %>% select(-Outcome)),
  y = clinical_stepwise_dtrain_80$Outcome,
  trControl = clinical_stepwise_train_control,
  tuneGrid = clinical_stepwise_tune_grid,
  metric= "ROC",
  method = "xgbTree"
)
registerDoSEQ()
clinical_stepwise_caret_model$bestTune
```

```{r clinical_stepwise care model save}
saveRDS(clinical_stepwise_caret_model, file="clinical_stepwise_caret_train.model")
```

#### Model results

#### AUC

```{r clinical_stepwise auc, warning=T, echo=FALSE, cache = T}
clinical_stepwise_pred_prob <- predict(clinical_stepwise_caret_model,
                                as.matrix(clinical_stepwise_dtest_20 %>% select(-Outcome)), type="prob")
#plot((roc(response=clinical_stepwise_dtest_20$Outcome, predictor=clinical_stepwise_pred$DEATH)),print.thres="best", print.thres.best.method="closest.topleft")
roc(response=clinical_stepwise_dtest_20$Outcome, predictor=clinical_stepwise_pred_prob$DEATH, levels = c("ALIVE", "DEATH"), direction = c("<"))
```

##### Predictivy values
```{r clinical_stepwise predictive values, cache = T}
clinical_stepwise_pred <- predict(clinical_stepwise_caret_model,
                           as.matrix(clinical_stepwise_dtest_20 %>% select(-Outcome)))
confusionMatrix(data = clinical_stepwise_pred,
                reference = clinical_stepwise_dtest_20$Outcome,
                positive = "DEATH")
```

Confidence intervals of Sensitivity and Specificity (Note that Outcome + means 'DEATH')

```{r clinical_stepwise predictive values w CIs, cache = T}
library(epiR)
options(digits = 3)
confusionMatrix(data = factor(clinical_stepwise_pred,levels=c("DEATH", "ALIVE")) , reference = factor(clinical_stepwise_dtest_20$Outcome, levels=c("DEATH", "ALIVE")), positive = "DEATH", mode=c("sens_spec"))$table -> clinical_stepwise_matrix_table
epiR::epi.tests(clinical_stepwise_matrix_table) -> clinical_stepwise_epitests

Sensitivity <- c(clinical_stepwise_epitests$elements$se, clinical_stepwise_epitests$elements$se.low, clinical_stepwise_epitests$elements$se.up)
Specificity <- c(clinical_stepwise_epitests$elements$sp, clinical_stepwise_epitests$elements$sp.low, clinical_stepwise_epitests$elements$sp.up)
`Positive Predictive Value` <- c(clinical_stepwise_epitests$elements$ppv, clinical_stepwise_epitests$elements$ppv.low, clinical_stepwise_epitests$elements$ppv.up)
`Negative Predictive Value` <- c(clinical_stepwise_epitests$elements$npv, clinical_stepwise_epitests$elements$npv.low, clinical_stepwise_epitests$elements$npv.up)
clinical_stepwise_epitests_df <- as.data.frame(rbind(Sensitivity, Specificity, `Positive Predictive Value`, `Negative Predictive Value`))
colnames(clinical_stepwise_epitests_df) <- c("Estimate", "Lower Interval", "Upper Interval")
knitr::kable(clinical_stepwise_epitests_df)
```

##### Variables importance

```{r clinical_stepwise variable importance, echo=FALSE, cache=T}
clinical_stepwise_var_importance <- varImp(clinical_stepwise_caret_model, scale = FALSE)
plot(clinical_stepwise_var_importance)
```

### Full Clinical model

Independet variables:

select_columns = c("Diabetes1", "PrimarySignCough1", "ChronicKidneyDisease1", "Respiratorydistress1"), 

* `Age`
* `SEX1`
* `PrimarySignFever1`
* `PrimarySignMuscularpain`
* `PrimarySignCough1`
* `PrimarySignDecreasedconsciousnes`
* `Respiratorydistress1`
* `Cancer1`
* `HIVAIDS1`
* `Asthma1`
* `ChrinicNeuorologicdisorder1`
* `Cardiovasculardisease1`
* `BloodChronicDisease1`
* `ChronicKidneyDisease1`
* `Diabetes1`
* `OtherLungDisease1`
* `contacthistoryCovid191`
* `Totalcommorbidities1`

#### Loading data

```{r full_clinical load  data, cache = TRUE}
dtrain_80 %>%
  select(Outcome,
         Age,
          SEX1,
          PrimarySignFever1,
          PrimarySignMuscularpain,
          PrimarySignCough1,
          PrimarySignDecreasedconsciousnes,
          Respiratorydistress1,
          Cancer1,
          HIVAIDS1,
          Asthma1,
          ChrinicNeuorologicdisorder1,
          Cardiovasculardisease1,
          BloodChronicDisease1,
          ChronicKidneyDisease1,
          Diabetes1,
          OtherLungDisease1,
          contacthistoryCovid191,
          Totalcommorbidities1) -> full_clinical_dtrain_80

dtest_20 %>%
  select(Outcome,
         Age,
          SEX1,
          PrimarySignFever1,
          PrimarySignMuscularpain,
          PrimarySignCough1,
          PrimarySignDecreasedconsciousnes,
          Respiratorydistress1,
          Cancer1,
          HIVAIDS1,
          Asthma1,
          ChrinicNeuorologicdisorder1,
          Cardiovasculardisease1,
          BloodChronicDisease1,
          ChronicKidneyDisease1,
          Diabetes1,
          OtherLungDisease1,
          contacthistoryCovid191,
          Totalcommorbidities1) -> full_clinical_dtest_20

```

#### Model train

```{r full_clinical model train, cache = TRUE}
library(doParallel)
cl <- makePSOCKcluster(detectCores())
registerDoParallel(cl)
full_clinical_train_control = trainControl(
  method = "cv",
  number = 5,
  verboseIter = FALSE,
  returnData = FALSE,
  returnResamp = "all", # save losses across all models
  classProbs = TRUE, # set to TRUE for AUC to be computed
  summaryFunction = twoClassSummary,
  allowParallel = TRUE
)

full_clinical_tune_grid = expand.grid(
  nrounds = 5000,
  max_depth = c(2, 4, 6, 8, 10),
  eta=c(0.5, 0.3, 0.1, 0.07),
  gamma = c(0, 0.01),
  colsample_bytree=c(1, 0.5),
  min_child_weight=1,
  subsample=c(1, 0.5)
)

# Recode https://stackoverflow.com/a/44131491/9664743
set.seed(0)
full_clinical_caret_model = train(
  x = as.matrix(full_clinical_dtrain_80 %>% select(-Outcome)),
  y = full_clinical_dtrain_80$Outcome,
  trControl = full_clinical_train_control,
  tuneGrid = full_clinical_tune_grid,
  metric= "ROC",
  method = "xgbTree"
)
registerDoSEQ()
full_clinical_caret_model$bestTune
```

```{r full_clinical care model save}
saveRDS(full_clinical_caret_model, file="full_clinical_caret_train.model")
```


#### Model results

##### AUC

```{r full_clinical auc, warning=T, echo=FALSE, cache = T}
full_clinical_pred_prob <- predict(full_clinical_caret_model,
                                as.matrix(full_clinical_dtest_20 %>% select(-Outcome)), type="prob")
#plot((roc(response=full_clinical_dtest_20$Outcome, predictor=full_clinical_pred$DEATH)),print.thres="best", print.thres.best.method="closest.topleft")
roc(response=full_clinical_dtest_20$Outcome, predictor=full_clinical_pred_prob$DEATH, levels = c("ALIVE", "DEATH"), direction = c("<"))
```

##### Predictivy values
```{r full_clinical predictive values, cache = T}
full_clinical_pred <- predict(full_clinical_caret_model,
                           as.matrix(full_clinical_dtest_20 %>% select(-Outcome)))
confusionMatrix(data = full_clinical_pred,
                reference = full_clinical_dtest_20$Outcome,
                positive = "DEATH")
```

Confidence intervals of Sensitivity and Specificity (Note that Outcome + means 'DEATH')

```{r full_clinical predictive values w CIs, cache = T}
library(epiR)
options(digits = 3)
confusionMatrix(data = factor(full_clinical_pred,levels=c("DEATH", "ALIVE")) , reference = factor(full_clinical_dtest_20$Outcome, levels=c("DEATH", "ALIVE")), positive = "DEATH", mode=c("sens_spec"))$table -> full_clinical_matrix_table
epiR::epi.tests(full_clinical_matrix_table) -> full_clinical_epitests

Sensitivity <- c(full_clinical_epitests$elements$se, full_clinical_epitests$elements$se.low, full_clinical_epitests$elements$se.up)
Specificity <- c(full_clinical_epitests$elements$sp, full_clinical_epitests$elements$sp.low, full_clinical_epitests$elements$sp.up)
`Positive Predictive Value` <- c(full_clinical_epitests$elements$ppv, full_clinical_epitests$elements$ppv.low, full_clinical_epitests$elements$ppv.up)
`Negative Predictive Value` <- c(full_clinical_epitests$elements$npv, full_clinical_epitests$elements$npv.low, full_clinical_epitests$elements$npv.up)
full_clinical_epitests_df <- as.data.frame(rbind(Sensitivity, Specificity, `Positive Predictive Value`, `Negative Predictive Value`))
colnames(full_clinical_epitests_df) <- c("Estimate", "Lower Interval", "Upper Interval")
knitr::kable(full_clinical_epitests_df)
```

##### Variables importance

```{r full_clinical variable importance, echo=FALSE, cache=T}
full_clinical_var_importance <- varImp(full_clinical_caret_model, scale = FALSE)
plot(full_clinical_var_importance)
```
